Reduced growth due to belowground sink limitation is not fully explained by reduced photosynthesis

### 2017-02-02 manuscript/editor_response_tree_physiology.Rmd
edits from editor to manuscript and editor response.  double checked the Amax to G stats

@@

@@

@@

@@

\
*Response*: 
\
- *Comment: 27. L483: There is no reduction in leaf photosynthesis due to the restricted soil volume, according to the data you reported in Table 3. Make clear what is leaf An. The reduction in whole plant assimilation may mainly be due to a reduction in plant leaf area, which in fact is only evident for the largest soil volume. Final plant leaf area can be added in table 1 with stats. It will be helpful as it is not so easy to visualize differences on figure 1.*

>>>>

\
*Response*: Editorial change made.  We have included more examples of C losses beyond respiration.  The rest of the paragraph is also adjusted to discuss "C loss" as a more general inclusive term.

@@

We have removed the statement "Leaf respiration is included in the calculation of *P*." and added a suffice description of how dark respiration was calculated for each 15 minute interval to then caluclate *A*~n~.

@@

\
*Response*: We agree with the editor here.  We have introduced a new term *A*~mass to represent *A*max on a mass basis.  This should help easily distinguish between the two variables in the text and is now consistent with Figure 4.

@@

\
*Response*: This is entirely true and was overlooked by the authors.  This statement has been removed from the results section. Perhaps this provides additionally evidence of how A and G can become decoupled under sink manipulation.

@@

\
*Response*: We have corrected this statment in the discussion to more accurately reflect the observed results.
\
- *Comment: 27. L483: There is no reduction in leaf photosynthesis due to the restricted soil volume, according to the data you reported in Table 3. Make clear what is leaf An. The reduction in whole plant assimilation may mainly be due to a reduction in plant leaf area, which in fact is only evident for the largest soil volume. Final plant leaf area can be added in table 1 with stats. It will be helpful as it is not so easy to visualize differences on figure 1.*

### 2017-02-02 manuscript/manuscript_pve_editor.Rmd
edits from editor to manuscript and editor response.  double checked the Amax to G stats

\
Most current models of plant growth are based on the physical principle of mass balance: C entering the plant through net photosynthesis must be balanced by C loss to respiration and C allocation to growth and storage. Plant growth is thus predicted as the outcome of the processes of photosynthesis, respiration and allocation. It is easy to see how source-limited growth can be implemented in such models: increasing source activity increases photosynthetic C uptake, flowing through to increased growth. It is less easy to know how to implement sink limitation of growth in such models, because it requires that we know whether, and how, sink limitation influences the contributing processes of photosynthesis, respiration and allocation. It has been suggested that sink limitation affects the C balance by causing a down-regulation of photosynthetic rate, possibly mediated by carbohydrate accumulation [@arp1991effects; @paul2001sink;@ainsworth2004testing; @ronchi2006growth; @nebauer2011photosynthesis], but there is also evidence that photosynthesis and growth can become strongly decoupled under sink limitation [@muller2011water;@leuzinger2013sink], suggesting that other processes must also be affected.

@@

where *L* is modelled total plant leaf area (m^2^), *P* is the daily C assimilation per unit leaf area (g C m^-2^ d^-1^, explained below), $\sigma$ is a self-shading parameter (explained below), and *R* is the mass based leaf respiration (g C d^-1^). Leaf respiration is included in the calculation of *P*.

\
*P* was predicted by using a coupled photosynthesis - stomatal conductance model [@farquhar1980biochemical;@medlyn2011reconciling] with the 'plantecophys' package [@duursma2015] in R with the mean photosynthetic parameters (*J*~max~, *V*~cmax~, *R*~dark~ and *g*~1~) for each treatment and meteorological data from an onsite weather station. Values of *J*~max~ and *V*~cmax~ are mean estimates from AC~i~ curves over two measurement periods (explained above), *R*~dark~ was empirically measured and the *g*~1~ parameter was generated by fitting the optimal stomatal conductance model from [@medlyn2011reconciling] to observed *g*~s~ values. Methods of the coupled leaf gas exchange model are also described in Duursma *et al.* [-@duursma2014peaked]. Combined with the meteorological parameters; PPFD, air temperature, and relative humidity, at 15 min intervals, leaf *A*~n~ rates ($\mu$mol CO~2~ m^-2^ s^-1^) were then predicted for each soil volume treatment. *P* was calculated by converting predicted rates to mass C gain over 15 min time steps (g m^-2^) and then summed for 24 h. This resulted in 120 unique values of *P* for each soil volume treatment, one value for each day of the experiment. Thus, each daily time step for model runs included a value of *P* that represented both treatment specific photosynthetic parameters and meteorological constraints across the duration of the experiment.

@@

##Gas exchange and photosynthetic parameters
At the first measurement campaign, both *A*~sat~ and *A*~max~ were significantly higher in the Free seedling treatment compared to seedlings in containers (both *P* < 0.001). Across all measurement campaigns, mean *A*~sat~ and *A*~max~ (Table `r pvetab3`) were consistently higher in Free seedlings than in containers (26 % and 29 %, respectively). The relationship between leaf starch and leaf N~f~ on a mass basis was marginally significant (*P* = 0.058), but *A*~max~ on a mass basis was correlated with both leaf N~f~ and leaf starch (both *P* < 0.001). We used predictions from the linear mixed effect model equation to visualize these relationships of *A*~max~ to either leaf N~f~ or leaf starch at multiple bin levels (n=5) of the co-variate parameter (Figure `r achemfig`). Across all measurement campaigns and treatments *A*~max~ was higher when leaf N~f~ was also higher, usually associated with low levels of leaf starch (Figure `r achemfig`a). *A*~max~  was also lower when leaf starch was high as higher leaf N~f~ often did not coincide with high leaf starch (Figure `r achemfig`b). Overall, *A*~max~ was positively correlated with final harvest biomass across all seedlings (*P* < 0.001).

@@

\
The lack of detected shifts in partitioning to fine roots provides evidence against an optimal foraging strategy for seedlings in containers. Root exudation may have increased with reduced rooting volumes to facilitate N uptake in favor of increasing partitioning to root biomass. The sensitivity of roots to their own exudates near obstructions may also prevent further growth [@semchenko2008foraging]. Here, we found that fine root length density was highest in smallest containers suggesting that root restriction likely occurred as simple function of available rooting space. Additionally, physical restriction of root proliferation could have impacted root development and morphology prior to shifts in mass partitioning. Here, increases in specific root length were detected in several of the soil volume treatments. This is not surprising as plants in containers have been shown to have different root morphology to field grown plants [@nesmith1998effect]. The poor soil quality used in our experiment and root restriction, however, likely decreased the capacity of this morphological response to increase N uptake.

>>>>

\
Most current models of plant growth are based on the physical principle of mass balance: C entering the plant through net photosynthesis must be balanced by C allocation to growth and storage and C losses such as respiration, root exudation, volatile organic compounds and tissue turnover. Plant growth is thus predicted as the outcome of the processes of photosynthesis, C allocation and C loss. It is easy to see how source-limited growth can be implemented in such models: increasing source activity increases photosynthetic C uptake, flowing through to increased growth. It is less easy to know how to implement sink limitation of growth in such models, because it requires that we know whether, and how, sink limitation influences the contributing processes of photosynthesis, C allocation and C loss. It has been suggested that sink limitation affects the C balance by causing a down-regulation of photosynthetic rate, possibly mediated by carbohydrate accumulation [@arp1991effects; @paul2001sink;@ainsworth2004testing; @ronchi2006growth; @nebauer2011photosynthesis], but there is also evidence that photosynthesis and growth can become strongly decoupled under sink limitation [@muller2011water;@leuzinger2013sink], suggesting that other processes must also be affected.

@@

where *L* is modelled total plant leaf area (m^2^), *P* is the daily C assimilation per unit leaf area (g C m^-2^ d^-1^, explained below), $\sigma$ is a self-shading parameter (explained below), and *R* is the mass based leaf respiration (g C d^-1^).

\
*P* was predicted by using a coupled photosynthesis - stomatal conductance model [@farquhar1980biochemical;@medlyn2011reconciling] with the 'plantecophys' package [@duursma2015] in R with the mean photosynthetic parameters (*J*~max~, *V*~cmax~, *R*~dark~ and *g*~1~) for each treatment and meteorological data from an onsite weather station. Values of *J*~max~ and *V*~cmax~ are mean estimates from AC~i~ curves over two measurement periods (explained above). *R*~dark~ was was estimated using Eq. (2) from ***Tjoekler 2008***, which was parameterized with a mean reference empircal measurement for each treatment, a Q~10~ value of 1.86 from a seperate experiment using the same tube stock seedlings and air temperature at each time step.

\
(2)
$$R~dark~ = R_{ref}Q_{10}~^{(T-T_{ref})/10}$$

The *g*~1~ parameter was generated by fitting the optimal stomatal conductance model from [@medlyn2011reconciling] to observed *g*~s~ values. Methods of the coupled leaf gas exchange model are also described in Duursma *et al.* [-@duursma2014peaked]. Combined with the meteorological parameters; PPFD, air temperature, and relative humidity, at 15 min intervals, leaf *A*~n~ rates ($\mu$mol CO~2~ m^-2^ s^-1^) were then predicted for each soil volume treatment. *P* was calculated by converting predicted rates to mass C gain over 15 min time steps (g m^-2^) and then summed for 24 h. This resulted in 120 unique values of *P* for each soil volume treatment, one value for each day of the experiment. Thus, each daily time step for model runs included a value of *P* that represented both treatment specific photosynthetic parameters and meteorological constraints across the duration of the experiment.

@@

##Gas exchange and photosynthetic parameters
At the first measurement campaign, both *A*~sat~ and *A*~max~ were significantly higher in the Free seedling treatment compared to seedlings in containers (both *P* < 0.001). Across all measurement campaigns, mean *A*~sat~ and *A*~max~ (Table `r pvetab3`) were consistently higher in Free seedlings than in containers (26 % and 29 %, respectively). The relationship between leaf starch and leaf N~f~ on a mass basis was marginally significant (*P* = 0.058), but *A*~max~ on a mass basis (*A*~mass~) was correlated with both leaf N~f~ and leaf starch (both *P* < 0.001). We used predictions from the linear mixed effect model equation to visualize these relationships of *A*~mass~ to either leaf N~f~ or leaf starch at multiple bin levels (n=5) of the co-variate parameter (Figure `r achemfig`). Across all measurement campaigns and treatments *A*~mass~ was higher when leaf N~f~ was also higher, usually associated with low levels of leaf starch (Figure `r achemfig`a). *A*~mass~  was also lower when leaf starch was high as higher leaf N~f~ often did not coincide with high leaf starch (Figure `r achemfig`b). Overall, *A*~max~ was positively correlated with final harvest biomass across all seedlings (*P* < 0.001).

@@

\
The lack of detected shifts in partitioning to fine roots provides evidence against an optimal foraging strategy for seedlings in containers. Root exudation may have increased with reduced rooting volumes to facilitate N uptake in favor of increasing partitioning to root biomass. The sensitivity of roots to their own exudates near obstructions may also prevent further growth [@semchenko2008foraging]. Here, we found that fine root length density was highest in smallest containers suggesting that root restriction likely occurred as simple function of available rooting space. Additionally, physical restriction of root proliferation could have impacted root development and morphology prior to shifts in mass partitioning. Although plants in containers have been shown to have different root morphology to field grown plants [@nesmith1998effect], increases in specific root length were detected in only one of the the soil volume treatments. The poor soil quality used in our experiment and root restriction, however, likely decreased the capacity of this morphological response to increase N uptake.
